---
title: Flight Map with ggplot2
author: Imran Kocabiyik
date: '2018-05-09'
slug: flight-map-with-ggplot2
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Flight Map with ggplot2

A flight map with ggplot2:  
  
  We need:  
  - A cartesian space. x $\mapsto$ `lat`, y $\mapsto$ `lon`  
- Line segments to represent routes  
- Polygons to represent to country borders  

## 1. Routes
Create some routes:  
```{r routes-df, message=FALSE, warning=FALSE}
# create routes
library(dplyr)
routes <- tribble(
  ~origin_airport, ~dest_airport,
  "IST", "TXL",
  "TXL", "TSF",
  "TXL", "ZRH",
  "TXL", "CPH",
  "TXL", "DLM",
  "TXL", "ADA",
  "TXL", "BOJ",
  "TXL", "TFS",
  "TXL", "BOS",
  "TXL", "FAE",
  "TXL", "CUZ",
  "TXL", "HKG"
)
```

## 2. Airport Locations

Add 4 more column: `origin_airport_lat`, `origin_airport_lon`, `destination_airport_lat`, `destination_airport_lon`   
```{r message=FALSE, warning=FALSE}
# read airport locations
library(readr)
library(dplyr)
airports <- read_csv("https://raw.githubusercontent.com/jpatokal/openflights/master/data/airports.dat", col_names = FALSE) %>% 
  select(X5, X7, X8) %>%
  rename(airport_code = X5,
         airport_lat = X7,
         airport_lon = X8)

routes <- routes %>%
  left_join(airports, by = c("origin_airport" = "airport_code")) %>% 
  rename(origin_airport_lat = airport_lat,
         origin_airport_lon = airport_lon) %>% 
  left_join(airports, by = c("dest_airport" = "airport_code")) %>% 
  rename(dest_airport_lat = airport_lat,
         dest_airport_lon = airport_lon)
routes
```


## 3. Visualize

Draw a basic plot:  
```{r airports}
# cartesian
library(ggplot2)
ggplot(data = routes) + 
  geom_segment(aes(x = origin_airport_lon, y = origin_airport_lat,
                   xend = dest_airport_lon, yend = dest_airport_lat)) +
  geom_text(aes(x = dest_airport_lon, y = dest_airport_lat, label = dest_airport)) +
  coord_equal() +
  xlab("Lon") +
  ylab("Lat")
```

It does not look like a flight map.  
We just draw line segments between `origin_airport` and `destination_airport`.  Next step is adding polygons. We can get the polygon data with `ggplot2::map_data` function.  

We can also change the theme by adding another layer: `theme_void`  


```{r flightmap}
library(ggplot2)
world <-  map_data("world") %>% tbl_df()
ggplot(data = routes) + 
  geom_polygon(data = world, aes(x = long, y = lat, group=group), 
               size = 0.2, color="white", alpha = 0.9, fill="lightgray") + 
  geom_segment(aes(x = origin_airport_lon, y = origin_airport_lat,
                   xend = dest_airport_lon, yend = dest_airport_lat),
               color = "brown") +
  geom_text(aes(x = dest_airport_lon, y = dest_airport_lat, label = dest_airport), size = 3) +
  theme_void()
```


We will have a more fancy plot by adding a `coord_map` layer.  
```{r flightmap2}
library(maps)
ggplot(data = routes) + 
  geom_polygon(data = world, aes(x = long, y = lat, group=group), 
               size = 0.2, color="white", alpha = 0.9, fill="lightgray") + 
  geom_segment(aes(x = origin_airport_lon, y = origin_airport_lat,
                   xend = dest_airport_lon, yend = dest_airport_lat),
               color = "brown") +
  coord_map("azequalarea", orientation = c(0, -10, 0)) +
  theme_void()
```
